---
title: "Lab 1"
author: Jackson Asiatico
date: "January 19, 2024"
output: html_document
---


# 2. Get the Data:

## Inspect the Data
* Are the column headers included in the file?
  - Yes. The column headers are the first row.
* How are data columns separated?
  - Each column is separated by commas.
* Are there comments in the data?
  - There does not appear to be any comments in the data.

## Load the Data
```{r read}
data <- read.table("lab1.dat", header = TRUE, sep = ",", comment.char = "")
```


# 3. Peruse the Data
## Print the names of the variables in the data table
```{r columnNames}
colnames(data)
```

## Global Summary
```{r summary}
summary(data)
```

# 4. Investigate the WineType Variable
## List values of WineType variable
```{r WineType}
WineType <- data$WineType
print(WineType)
unique(data$WineType)
```

## Compute the levels of that factor
```{r WineTypeLevels}
WineTypeLevels <- levels(factor(WineType))
print(WineTypeLevels)
```

## Compute the number of records of each level
```{r NumWineType}
NumWineType <- table(WineType)
print(NumWineType)
```

# 5. Investigate the ColorIntensity variable
* Compute the mean of the ColorIntensity variable
```{r ColorIntense}
ColorIntenseMean <- mean(data$ColorIntensity)
cat("The mean of Color Intensity is: ", ColorIntenseMean)
```

* Compute variance of the ColorIntensity variable
```{r ColorIntense1}
ColorIntenseVar <- var(data$ColorIntensity)
cat("The variance of Color Intensity is: ", ColorIntenseVar)
```

* Compute standard deviation of the ColorIntensity variable
```{r ColorIntense2}
ColorIntenseSTD <- sd(data$ColorIntensity)
cat("The standard deviation of Color Intensity is: ", ColorIntenseSTD)
```

* Compute 95% confidence interval of the ColorIntensity variable
```{r ColorIntense3}
ColorIntenseConfInter <- t.test(data$ColorIntensity)
print(ColorIntenseConfInter)
```


# 6. Investigate the Hue variable for WineType Class1
* Compute variance of the Hue variable across only the records with WineType=Class1
```{r HueClass1}
Class1HueVar <- var(subset(data$Hue, data$WineType == "Class1"))
cat("The variance of the Hue variable for Class1: ", Class1HueVar)
```

* Compute standard deviation of the Hue variable across only the records with WineType=Class1
```{r HueClass12}
Class1HueSTD <- sd(subset(data$Hue, data$WineType == "Class1"))
cat("The standard deviation of the Hue variable for Class1: ", Class1HueSTD)
```

* Compute 95% confidence interval of the Hue variable across only the records with WineType=Class1
```{r HueClass13}
Class1HueConf <- t.test(subset(data$Hue, data$WineType == "Class1"))$conf.int
cat("The 95% confidence interval of the Hue variable for Class1: ","\n", Class1HueConf)
```

# 7. Perform simple math operations

* Subtract 120 from 536
```{r SimpleMath1}
536-120
```

* Multiply 3 and 7
```{r SimpleMath2}
3*7
```

* Divide 12 by 3
```{r SimpleMath3}
12/3
```

* Take the base 2 logarithm of 128
```{r SimpleMath4}
log2(128)
```

* Compute nine cubed
```{r SimpleMath5}
9^3
```

# 8. Perform math operations on data columns

* Compute the vector resulting from a pair-wise multiplication of all Hue variable values with ColorIntensity values
```{r MathOps}
A = data$Hue * data$ColorIntensity
print(A)
```
* Sum the resulting vector from the previous step
```{r MathOps1}
B = sum(A)
print(B)
```

# 9. Drawing random numbers

* Draw 10 numbers from the uniform distribution, between 0.0 and 10.0
```{r RandNum1}
RandUnif <- runif(10, min=0.0,max=10.0)
print(RandUnif)
```

* Draw 10 numbers from the Normal distribution, mean of 1.0, standard deviation of 0.3
```{r RandNum2}
RandNorm <- rnorm(10,mean=1.0,sd=0.3)
print(RandNorm)
```

* Compute PDF result for the standard Normal distribution at the value 0.12
```{r RandNum3}
PDFNorm <- dnorm(0.12)
print(PDFNorm)
```
* Compute the probability that a value drawn from the standard Normal distribution will be less than 1.96
```{r RandNum4}
Prob <- pnorm(1.96)
print(Prob)
```

# 10. Simulation
* Perform a Monte Carlo sampling simulation of customers arriving at a queue over the period of 10 minutes, assuming this process is governed by the Poisson distribution and that 2 customers arrive per minute on average 

```{r Sim1}
CustPerMin <- 2   # Customers arriving per minute
TPeriod <- 10     # Time period
x <- rpois(TPeriod, lambda=CustPerMin)
print(x)
```

* Compute the total number of people standing in line at each minute for one such simulation?
```{r Sim2}
CustInLine <- cumsum(x)
print(CustInLine)
```

* Enter the R code below, which defines a function. Call that function using n=10, then again using n=10000, then again using n=10000000. What does the function estimate?
```{r Sim3}
estimate <- function(n) {
  return (4*sum((runif(n)^2 + runif(n)^2) < 1)/n)
}

Res1 <- estimate(10)
print(Res1)

Res2 <- estimate(10000)
print(Res2)

Res3 <- estimate(10000000)
print(Res3)
```